Old or Heavy?\\Decaying Gracefully with Age/Weight Shapes.

### 2019-02-27 papers/CADE19/paper.tex
intro

\documentclass{llncs}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage[dvipsnames]{xcolor}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\title{Old or Heavy?\\Decaying Gracefully with Age/Weight Shapes.}
\titlerunning{Dynamic Strategy Priority}
\authorrunning{Reger \and Rawson}
\author{Michael Rawson \and Giles Reger}
\institute{University of Manchester, Manchester, UK}

@@

\section{Introduction}
\label{sec:introduction}
\todo{standard Vampire~\cite{vampire} fluff here}.\\
\todo{explain a generic given-clause algorithm, define clause age and weight --- and how the ``balancing'' algorithm works internally}.\\
\todo{remark on portfolio modes --- more choice is (sometimes) a good thing}.

\section{Optimising Age/Weight Ratios}
\subsection{Fixed AWR}
Changing Vampire's fixed AWR has a significant effect on the number of clauses required to be processed before a proof is found.
For example, Figure \ref{fig:example-optimal-awr} shows the effect of varying AWR on an illustrative problem from TPTP: a smaller number of active clauses means that fewer clauses were processed, which in general means that a proof was found faster\footnote{
It should be noted that if a small number of clauses are extremely expensive to process it may be slower than a larger number of less-expensive clauses, but in general this is a good heuristic measure for prover performance.
It also avoids reproducibility issues involved with using system timing approaches.
}.
On this problem, a good AWR value is over 400\% better by this metric than the worst AWR value.

@@

\section{Variable AWR for Vampire}
If changing AWR is sometimes useful for finding proofs more quickly, this could be useful as a Vampire option.
In general we would like to describe any possible sequence that the AWR could follow during proof search.
However, some details constrain the design space:
\begin{enumerate}
	\item Changing the AWR too frequently or sharply has little effect, due to the ``balancing'' algorithm --- see Section \ref{sec:introduction}.
	\item A general (configurable) \emph{shape} is more likely to be widely applicable than a specific series of data points.
	\item The shape must extend naturally to an indefinitely-long proof search.
\end{enumerate}

@@

\section{Future Work}
\todo{more shapes, better ways of doing frequency decay, integrate into existing strategy schedules}.

\section{Conclusions}
\todo{this is a thing, you can do it, you should do it (?), we can probably do better}.

>>>>

\documentclass{llncs}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage[dvipsnames]{xcolor}
\usepackage{amssymb}
\usepackage{amsmath}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

\title{Old or Heavy?\\Decaying Gracefully with Age/Weight Shapes.}
\titlerunning{Dynamic Strategy Priority}
\authorrunning{Reger \and Rawson}
\author{Michael Rawson \and Giles Reger}
\institute{University of Manchester, Manchester, UK}

%
%   Logical notation
%
\newcommand{\imply}{\rightarrow}
%\renewcommand{\implies}{\imply}
\newcommand{\liff}{\Leftrightarrow}
\newcommand{\lniff}{\not\Leftrightarrow}
\newcommand{\iffl}{\leftrightarrow}
\newcommand{\xor}{\otimes}
\newcommand{\orl}{\vee}
\newcommand{\bigorl}{\bigvee}
\newcommand{\andl}{\wedge}
\newcommand{\bigandl}{\bigwedge}
\newcommand{\notl}{\neg}
\renewcommand{\models}{\vDash}
\newcommand{\nmodels}{\nvDash}
\newcommand{\subst}[2]{#1 \mapsto #2}       % for substitutions
\newcommand{\Subst}[1]{\{#1\}}      % for substitutions
\newcommand{\emptysubst}{\varepsilon}
\newcommand{\true}{\mathsf{true}}      % boolean value true
\newcommand{\false}{\mathsf{false}}     % boolean value false
\newcommand{\eql}{\simeq}           % equality
\newcommand{\neql}{\not\simeq}           % equality

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}

At the core of all modern saturation-based theorem provers is \emph{clause selection}. It has been known for a long time that the way that clauses are selected in the given clause algorithm is important and early work in 80s already introduce the concept of selecting clauses as a ratio between the age and weight of a clause with the intuition that this balances between breadth and depth first search of the clause space. Whilst complex

\cite{DBLP:conf/cade/SchulzM16}

We explore this

After covering relevant background material in Section~\ref{sec:background} the remainder of the paper makes the following contributions:
\begin{itemize}
	\item An experimental 
	\item 
	\item An implementation of this approach in the Vampire~\cite{vampire} theorem prover and an experimental analysis demonstrating its impact. This new method for clause selection led to X
\end{itemize}
Conclusions are drawn and future work discussed in Section~\ref{sec:conclusion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background} \label{sec:background}

This section introduces the relevant background for the rest of the paper.

%%%%%%
\paragraph{First-Order Logic.} We consider the standard first-order predicate logic with
equality. Terms are of the form $f(t_1,\ldots,t_n), c$ or $x$ where
$f$ is a \emph{function symbol} of arity $n \geq 1$, $t_1,\ldots, t_n$ are terms, $c$ is a zero arity function symbol (i.e. a constant) and $x$ is a variable. 
%
The \emph{weight} of a term $t$ is defined as $w(t) =1$ if $t$ is a variable or a constant
and as $w(t) = 1+\sum_{i=1,\ldots,n} w(t_i)$ if $t$ is of the form
$f(t_1,\ldots,t_n)$. In other words, the weight or a term is the
number of symbols in it.
%
Atoms are of the form $p(t_1,\ldots,t_n), q$ or $t_1 \eql t_2$ where $p$ is a \emph{predicate symbol} of arity $n$, $t_1, \ldots, t_n$ are terms, $q$ is a zero arity predicate symbol and $\eql$ is the \emph{equality symbol}. 
%
The weight function naturally extends to atoms: $w(p(t_1,\ldots,t_n)) = 1+\sum_{i=1,\ldots,n} w(t_i)$, $w(t_1 \eql t_2) = w(t_1)+w(t_2)$,
and $w(q)=1$.
%
A literal is either an atom $A$, in which case we call it \emph{positive}, or a negation $\notl A$,
in which case we call it \emph{negative}.
%We write negated equalities as $t_1 \neql t_2$. 
%
The weight of a literal is the weight of the corresponding atom. 
% 
%We write $t[s]_p$ and $L[s]_p$ to denote that a term $s$ occurs in a term $t$ (in a literal $L$) at a position $p$.
%
A \emph{clause} is a disjunction of literals $L_1 \orl \ldots \orl L_n$ for $n \geq 0$.
We disregard the order of literals and treat a clause as a multiset. 
%
The weight of a clause is the sum of the weight of its literals.

%%%%%%
\paragraph{The Given Clause Algorithm.}

Saturation-based theorem provers \emph{saturate} a set of clauses $S$ with respect to an inference system $\mathbb{I}$ e.g. they compute a set of clauses $S'$ by applying rules in $\mathbb{I}$ to clauses in $S$ until no new clauses are generated. 
%
If the empty clause is generated then $S$ is unsatisfiable.
%
Calculi such as resolution and superposition have conditions that ensure \emph{completeness}, which means that a saturated set $S$ is satisfiable if it does not contain the empty clause as an element.
%
As first-order logic is only semi-decidable, it is not necessarily the case that $S$ has a finite saturation, and even if it does it may be unachievable in practice using the available resources. 
%
Therefore, much effort in saturation-based first-order theorem proving involves 
controlling proof search to make finding the empty clause more likely (within reasonable resource bounds). 
%
One important notion is that of \emph{redundancy}, being able to remove clauses from the search space that are not required. 
%
Another important notion, the subject of this paper, are literal selections that place restrictions on the inferences that can be performed. 
%
Both notions come with additional requirements for completeness.

\todo{explain a generic given-clause algorithm; define clause age}

%%%%%%
\paragraph{Age-Weight Based Clause Selection.}

\todo{ how the ``balancing'' algorithm works internally}

%%%%%%
\paragraph{Portfolio Solvers.}

\todo{remark on portfolio modes --- more choice is (sometimes) a good thing}.

%%%%%%
\paragraph{Related Work.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Optimising Age/Weight Ratios}
\subsection{Fixed AWR}
Changing Vampire's fixed AWR has a significant effect on the number of clauses required to be processed before a proof is found.
For example, Figure \ref{fig:example-optimal-awr} shows the effect of varying AWR on an illustrative problem from TPTP: a smaller number of active clauses means that fewer clauses were processed, which in general means that a proof was found faster\footnote{
It should be noted that if a small number of clauses are extremely expensive to process it may be slower than a larger number of less-expensive clauses, but in general this is a good heuristic measure for prover performance.
It also avoids reproducibility issues involved with using system timing approaches.
}.
On this problem, a good AWR value is over 400\% better by this metric than the worst AWR value.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variable AWR for Vampire}
\label{sec:variable}

If changing AWR is sometimes useful for finding proofs more quickly, this could be useful as a Vampire option.
In general we would like to describe any possible sequence that the AWR could follow during proof search.
However, some details constrain the design space:
\begin{enumerate}
	\item Changing the AWR too frequently or sharply has little effect, due to the ``balancing'' algorithm --- see Section \ref{sec:introduction}.
	\item A general (configurable) \emph{shape} is more likely to be widely applicable than a specific series of data points.
	\item The shape must extend naturally to an indefinitely-long proof search.
\end{enumerate}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}
\label{sec:conclusions}

\todo{this is a thing, you can do it, you should do it (?), we can probably do better}.

\todo{more shapes, better ways of doing frequency decay, integrate into existing strategy schedules}.
